---
title: "Dataset"
author: "Matthew Hoctor"
date: "5/6/2022"
output:
  html_document:
    number_sections: no
    theme: lumen
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# library(dplyr)
# library(readxl)
library(openxlsx)
library(tidyverse)
library(janitor)
library(gtsummary)
# library(ggplot2)
# library(gridExtra)        #grid.arrange for multiple ggplots
# library(reshape2)       #melt function for simple longform datasets
# library(CarletonStats)
# library(pwr)
# library(BSDA)
# library(exact2x2)
# library(car)
# library(dvmisc)
# library(emmeans)
# library(DescTools)
# library(DiagrammeR)     #for plotting trees
# library(nlme)
# library(doBy)
# library(geepack)
# library(rje)
# library(ISLR2)
# library(psych)
# library(MASS)
# library(caret)            #for confusionMatrix function
# library(rje)
# library(class)          #for knn function
# library(e1071)          #for naiveBayes function & SVM svm() funcion
# library(boot)           #for boot function
# library(covTest)        #for covTest function
# library(leaps)          #for regsubsets function for best subset selection
# library(broom)
# library(glmnet)         #for glmnet() for shrinkage methods
# library(doParallel)     #for parallel computing in glmnet(); does not work
# library(pls)            #for pcr function
# library(qpcR)           #for RSS function
# library(splines)        #for bs() function for basis function of regression splines
# library(quantreg)       #for quantreg() for quantile regression
# library(tree)           #older package for tree regression
# library(rpart)          #more maintained package for tree regression
# library(rattle)         #for visually appealing tree plotting
# library(randomForest)   #for random forest technique
# library(party)          #?cforest function for plotting random forests?
# library(xgboost)
# library(gbm)            #more gradient boosting functions
# library(LiblineaR)
# library(svmpath)
# library(msvmpath)
# library(scatterplot3d)    #for the 3d scatterplot of pca
# library(mclust)           #for cluster analysis; also imputeData() & imputePairs() functions
# library(tightClust)       #another clustering library; tight.clust() function
# library(softImpute)       #for imputing missing data
# library(factoextra)       #for plotting clusters from k-means
# library(keras)
# library(neuralnet)          #traditional neutal network package
```

# Import Data

Import the main dataset

```{r}
data_v2 <- read.xlsx(xlsxFile = "data/OPTIONS DATABASE v2.0_HAV_HBV.xlsx")
```

Import the extra data

```{r}
allergy <- read.xlsx(xlsxFile = "data/HAV_HBV data_pull_de_ID.xlsx",
                      sheet = "Allergies")
# vaccines administered at any site
# NB Vaccines without an administration date were ordered but not given
vaccine <- read.xlsx(xlsxFile = "data/HAV_HBV data_pull_de_ID.xlsx",
                      sheet = "Any_vaccine",
                      detectDates = TRUE)
hep_tests <- read.xlsx(xlsxFile = "data/HAV_HBV data_pull_de_ID.xlsx",
                      sheet = "hepatitis_tests")
# Old labs data
labs <- read.xlsx(xlsxFile = "data/HAV_HBV data_pull_de_ID.xlsx",
                      sheet = "other_labs")

#From new data pull, includes day that the lab was taken
labs2 <- read.xlsx(xlsxFile = "data/Matt's data_de_ID_6_3_22.xlsx",
                      sheet = "CBC_CMP_labs")
vaccine2 <- read.xlsx(xlsxFile = "data/Matt's data_de_ID_6_3_22.xlsx",
                      sheet = "vaccines_during_admit")
icu <- read.xlsx(xlsxFile = "data/Matt's data_de_ID_6_3_22.xlsx",
                      sheet = "ICU_days")
```

Import chart review data

```{r}
chart <- read.xlsx(xlsxFile = "data/STI_chart_review - deID.xlsx",
                      sheet = "chart_review")
```

# Cleaning the dataset

## Initial cleaning

Creates a new dataset to modify:

```{r}
dat <- data_v2
```

Removes empty columns and sparse columns (columns with only one non-NA value):

```{r}
sparsity <- colSums(is.na(dat))
empty_columns <- sparsity == nrow(dat)
sparse_columns <- sparsity >= (nrow(dat) - 1)

# how many empty columns
summary(empty_columns)
# how many sparse columns
summary(sparse_columns)

# remove sparse columns
dat <- dat[, !sparse_columns]

# cleanup
rm(empty_columns, sparse_columns, sparsity)
```

## Clean Variable names

```{r}
# Replaces . with _
names(dat) <- str_replace_all(string = names(dat), 
                                  pattern =  "\\.",  
                                  replacement = "_")

# Removes all text after the colon; i.e. the unnecessary encoding info
names(dat) <- str_remove(string = names(dat), 
                             pattern =  "\\:.*")
```

## Widen infection date information

Creates a wide-format dataframe with Infection_ID in columns according to year of admission:

```{r}
dates <- dat %>% 
  select(Patient_ID, Infection_ID, Year_of_Admission) %>%
  pivot_wider(
    id_cols = Patient_ID,
    names_from = Year_of_Admission,
    values_from = Infection_ID,
    names_prefix = "i_",
    names_sort = TRUE
)
```

Creates a wide-format dataframe and finds the first (lowest value) infection:

```{r}
# creates a constant variable to pivot from
dat$constant <- "infections"
infections <- dat %>% 
  select(Patient_ID, Infection_ID, constant) %>%
  pivot_wider(
    id_cols = Patient_ID,
    names_from = constant,
    values_from = Infection_ID,
)

# finds the first infection
infections$first <- lapply(infections$infections, FUN = min)

# cleans up the constant variable:
dat$constant <- NULL
```

We can now merge these admission dates into the main dataset:

```{r}
dat <- merge(x = dat, 
             y = dates,
             by = "Patient_ID",
             all.x = TRUE)

dat <- merge(x = dat, 
             y = infections,
             by = "Patient_ID",
             all.x = TRUE)

# cleanup the temporary dataframes (dates, infections)
remove(list = c("dates", "infections"))
```

## Merge chart review data

```{r}
dat <- merge(x = dat,
             y = chart,
             by = c("Patient_ID", "Infection_ID", "age_at_admission"),
             all.x = TRUE)

# cleanup chart review dataframe
remove(chart)
```

## Widen & merge lab data

```{r}
# widen the lab data
labsW <- labs %>%
  filter(is.na(Lab_Component_Result_Value) == FALSE) %>%
  pivot_wider(id_cols = c("Patient_ID", "Infection_ID"),
              names_from = Lab_Component_Name,
              values_from = Lab_Component_Result_Value,
              values_fill = NA)
# merge the lab data
dat <- merge(x = dat,
             y = labsW,
             by = c("Patient_ID", "Infection_ID"),
             all.x = TRUE)
# cleanup
remove(labsW, labs)
```

## Widen & merge hepatitis test data

```{r}
# widen
hepW <- hep_tests %>%
  filter(is.na(Lab_Component_Result_Value) == FALSE) %>%
  pivot_wider(id_cols = c("Patient_ID", "Infection_ID"),
              names_from = Lab_Component_Name,
              values_from = Lab_Component_Result_Value,
              values_fill = NA,
              names_sort = TRUE)
# merge the hepatitis data
dat <- merge(x = dat,
             y = hepW,
             by = c("Patient_ID", "Infection_ID"),
             all.x = TRUE)
# cleanup
remove(hepW, hep_tests)
```

## Widen & merge vaccine data

Tabulating vaccines administered in-hospital:

```{r}
table(vaccine2$Generic_Name)
```

```{r}
# recode a brand name variable for differently named generic drugs
vaccine2$brand <- recode(vaccine2$Generic_Name,
                         "hepatitis A vaccine (PF) 1,440 ELISA unit/mL intramuscular syringe" = "Havrix",
                         "hepatitis A virus vaccine (PF) 1,440 ELISA unit/mL IM suspension" = "Havrix",
                         "hepatitis B vaccine 20 mcg/0.5 mL-adjuvant CpG 1018 (PF) IM solution" = "HeplisavB",
                         "hepatitis B vaccine 20 mcg/0.5 mL-adjuvant CpG 1018 (PF) IM syringe" = "HeplisavB",
                         "hepatitis B virus vaccine recomb (PF) 20 mcg/mL intramuscular susp" = "EngerixB")

# widen
vaccineW2 <- vaccine2 %>%
  filter(is.na(Generic_Name) == FALSE) %>%
  pivot_wider(
    id_cols =  c("Patient_ID", "Infection_ID"),
    names_from = brand,
    values_from = admin_date_admission_date,
    values_fill = NA,
    names_sort = TRUE
)

# convert NULL to NA
vaccineW2$EngerixB[vaccineW2$EngerixB == "NULL"] <- NA
vaccineW2$Havrix[vaccineW2$Havrix == "NULL"] <- NA
vaccineW2$HeplisavB[vaccineW2$HeplisavB == "NULL"] <- NA

# merge
dat <- merge(x = dat,
             y = vaccineW2,
             by = c("Patient_ID", "Infection_ID"),
             all.x = TRUE)

# cleanup
remove(vaccineW2)
```


## Widen & merge vaccine ordering/administration data



## Clean the full dataset

Re-cleaning the variable names

```{r}
# Replaces . with _
names(dat) <- str_replace_all(string = names(dat), 
                                  pattern =  "\\.",  
                                  replacement = "_")

# Removes all text after the colon; i.e. the unnecessary encoding info
names(dat) <- str_remove(string = names(dat), 
                             pattern =  "\\:.*")
```

Replacing NULL values with NA

```{r}
dat <- dat %>% replace(.=="NULL", NA)
```

# Exposure Variables

## HCV

```{r}
dat$hcv <- recode_factor(dat$`hcv_ab_or_pcr_positive`,
                         '0' = "HCV-",
                         '1' = "HCV+")
```

We can validate this data by cross-tabulation with chart-review data.  First we can find those patients with positive HCV Ab or HCV DNA:

```{r}
# Uncomment to tabulate HCV PCR & Ab results individually 
# table(dat$HCVAb_Result, useNA = "ifany")
# table(dat$HCV_PCR_result, useNA = "ifany")

dat$hcv2 <- NA
dat$hcv2[dat$HCVAb_Result == "Not Detected" | dat$HCVAb_Result == "Negative" | dat$HCVAb_Result == "Indeterminate; confirmation to follow"| dat$HCV_PCR_result == 0] <- 0
dat$hcv2[dat$HCVAb_Result == "Detected" | dat$HCVAb_Result == "Positive" | dat$HCV_PCR_result == 1] <- 1

table(dat$hcv2, dat$HCVAb_Result, useNA = "ifany")
table(dat$hcv2, dat$hcv, useNA = "ifany")
```

### Discordance

```{r}
table(dat$hcv2, dat$hcv, useNA = "ifany")
```

There is not complete concordance in this encoding.  We can also consider HCV status at admission, to see if this will rectify it:

```{r}
table(dat$HCV_status_at_admission, useNA = "ifany")
table(dat$HCV_status_at_admission, dat$hcv2, useNA = "ifany")
dat$hcv3 <- 0 # dat$hcv2
dat$hcv3[dat$HCV_status_at_admission != "No prior positive HCV test"] <- 1
table(dat$hcv2, dat$hcv3, useNA = "ifany")
table(dat$hcv, dat$hcv3, useNA = "ifany")
```

Now we can combine the HCV chart lab data with the main dataset

```{r}
dat$hcv4 <- dat$hcv

# incorporating the new HCV Ab & PCR info from the hcv2 variable:
dat$hcv4[is.na(dat$hcv4) & dat$hcv2 == 0] <- "HCV-"
dat$hcv4[dat$hcv2 == 1] <- "HCV+"
table(dat$hcv, useNA = "ifany")
table(dat$hcv2, dat$hcv, useNA = "ifany")
table(dat$hcv4, dat$hcv, useNA = "ifany")
table(dat$hcv4, useNA = "ifany")
```

We can now combine the HCV chart clinical status prior to admission data with the main dataset

```{r}
table(dat$hcv4, useNA = "ifany")
table(dat$hcv3, dat$hcv4, useNA = "ifany")
dat$hcv4[is.na(dat$hcv4) & dat$hcv3 == 0] <- "HCV-"
dat$hcv4[dat$hcv3 == 1] <- "HCV+"
table(dat$hcv3, dat$hcv, useNA = "ifany")
table(dat$hcv4, dat$hcv, useNA = "ifany")
table(dat$hcv4, useNA = "ifany")
```

This produces a combined HCV+ variable (a composite of the chart review data, HCV lab data, and HCV status prior to admission) with no missing values.

## Homeless

```{r}
dat$homeless <- recode_factor(dat$`housing`,
                       '0' = "Housed",
                       '1' = "Homeless")
```

## IVDU

```{r}
dat$ivdu <- 0
dat$ivdu[dat$opioid_used == 1] <- 1
dat$ivdu[dat$methamphetamine_used == 1] <-1
dat$ivdu[dat$cocaine_used == 1] <- 1
dat$ivdu <- factor(dat$ivdu,
                   levels = c(0,1),
                   labels = c("No Known IVDU", "IVDU"),
                   ordered = TRUE)
```

# Covariates

## Age

```{r}
dat$age <- dat$age_at_admission

summary(dat$age)
```

## Gender

```{r}
dat$gender <- recode_factor(dat$gender_m_or_f_lowercase,
                        'm' = 0,
                        'f' = 1)
dat$gender <- factor(dat$gender,
                     levels = c(0,1),
                     labels = c("Male", "Female"),
                     ordered = TRUE)

# gender variable for T1
dat$genderM <- ifelse(dat$gender == "Male", 1, 0)

summary(dat$gender)
```

## Race/Ethnicity

```{r}
dat$race_eth <- NA
dat$race_eth[dat$race == "white"] <-1
dat$race_eth[dat$race == "black"] <- 2
dat$race_eth[dat$ethnicity == "hispanic"] <-3
dat$race_eth[dat$race == "american indian/alaska native"] <- 4
dat$race_eth[dat$race == "asian"] <- 5
dat$race_eth[dat$race == "other pacific islander"] <- 6
dat$race_eth <- factor(dat$race_eth,
                       levels = 1:6,
                       labels = c("White", "Black", "Hispanic White", "AI/AN", "Asian", "Pacific Islander"))
table(dat$race_eth, useNA = "ifany")
```

Create collapsed reace/eth for T1:
  
```{r}
dat$race_eth_t1 <- 4
dat$race_eth_t1[dat$race == "white"] <-1
dat$race_eth_t1[dat$race == "black"] <- 2
dat$race_eth_t1[dat$ethnicity == "hispanic"] <-3
dat$race_eth_t1 <- factor(dat$race_eth_t1,
                          levels = 1:4,
                          labels = c("White", "Black", "Hispanic White", "AI/AN, Asian, Pacific Islander, or Unknown"))

table(dat$race_eth, dat$race_eth_t1, useNA = "ifany")
```

## Medicaid Eligibility

```{r}
dat$medicaid <- 0
dat$medicaid[dat$insurance == "medicaid"] <- 1
dat$medicaid[is.na(dat$insurance)] <- NA
dat$medicaid <- factor(dat$medicaid,
                       levels = 0:1,
                       labels = c("Non-Medicaid Insurance", "Medicaid"),
                       ordered = TRUE)

# create medicaid variable for table 1
dat$medicaid_t1 <- 0
dat$medicaid_t1[dat$medicaid == "Medicaid"] <- 1

table(dat$medicaid, dat$medicaid_t1, useNA = "ifany")
```

## HIV Status

```{r}
dat$hiv2 <- factor(dat$hiv,
                  levels = 0:2,
                  labels = c("HIV-", "HIV+", NA))

# Create HIV variable for Table 1
dat$hiv_t1 <- 0
dat$hiv_t1[dat$hiv2 == "HIV+"] <- 1

table(dat$hiv2, dat$hiv_t1)
```

We can cross-tabulate this with info from the chart review

```{r}
# reduces output to a readable length
dat$HIV_Result[dat$HIV_Result == "HIV-1 was not detected in this sample.  These results are most likely suggestive of either the absence of HIV viremia or the presence of low-level HIV viremia at or below the assays lower limit of detection."] <- "HIV-1 was not detected in this sample."

table(dat$HIV_Result, useNA = "ifany")
dat$hiv3 <- 0
dat$hiv3[is.na(dat$HIV_Result)] <- NA
dat$hiv3[dat$HIV_Result == "HIV-1 RNA was detected in this sample." | dat$HIV_Result == 4900 | dat$HIV_Result == 64 | dat$HIV_Result == 760] <- 1
table(dat$hiv3, dat$HIV_Result, useNA = "ifany")

table(dat$hiv3, dat$hiv2, useNA = "ifany")
```

Partial concordance.  This is possibly due to undetectable viral load during treatment.

```{r}
hiv_discordance_subset <- subset(dat, hiv2 == "HIV+" & hiv3 == 0)
hiv_discordance_subset$HIV_test
hiv_discordance_subset$HIV_Result
# cleanup
rm(hiv_discordance_subset)
```

Undetectable viral load on a quantitative PCR test is consistent with the above hypothesis that the discordant observation may be due to undetectable viral load during treatment.

## Allergies

Yeast allergy is a contraindication to all available HBV vaccines, with the exception of PreHevbrio; while aminoglycoside allergy is a contraindication only to Twinrix.  Of the reactions listed in the dataset, gentamycin is an aminoglycoside, with high crossreactivity to neomycin (the preservative in Twinrix), and the PCV-13 vaccine is also a yeast-derived vaccine, and could presumably cause cross reaction.

```{r}
allergyW <- pivot_wider(allergy,
                        id_cols = Patient_ID,
                        names_from = Allergen_Name,
                        values_from = Allergy_First_Reaction,
                        names_sort = TRUE)

#Deriving yeast allergy data from PCV-13 allergy:
allergyW$yeast_allergy <- ifelse(allergyW$`PNEUMOC 13-VAL CONJ-DIP CR(PF)` == "NULL", 0, 1)

#Deriving aminoglycoside allergy data from gentamycin allergy:
allergyW$ag_allergy <- ifelse(allergyW$GENTAMICIN == "NULL", 0, 1)
```

We can now examine how many individuals have each allergy

```{r}
sum(allergyW$ag_allergy)
sum(allergyW$yeast_allergy)
```

Only one of each.  We can now examine the reactions to each:

```{r}
allergyW[allergyW$yeast_allergy == 1,]$`PNEUMOC 13-VAL CONJ-DIP CR(PF)`
allergyW[allergyW$ag_allergy == 1,]$GENTAMICIN
```

The patient with 'allergy' to the PCV13 vaccine experienced pruritis (a common, >10%, reaction to this vaccine), whereas the patient with aminoglycoside allergy experienced hives.  An IgE mediated reaction to PCV-13 may be considered a contraindication to HBV vaccination, as common ingredients are shared, but pruritis does not qualify. Havrix contains neomycin, which may cross-react with gentamycin.  We can now join this data into the main dataset:

```{r}
dat <- merge(x = dat, 
             y = select(.data = allergyW,
                            Patient_ID, 
                            ag_allergy,
                            yeast_allergy),
             by = "Patient_ID",
             all.x = TRUE)

# cleanup
remove(allergyW, allergy)
```

## FIB-4 & liver biomarkers

We can find the FIB-4 initial values of the relevant labs (AST, ALT, platelets)

```{r}
dat$alt_f <- lapply(dat$`ALT (SGPT)`, FUN = first) %>% str_remove(pattern = "[:symbol:]") %>% as.numeric() 
dat$ast_f <- lapply(dat$`AST(SGOT)`, FUN = first) %>% str_remove(pattern = "[:symbol:]") %>% as.numeric() 
dat$plt_f <- lapply(dat$`PLATELET COUNT`, FUN = first) %>% str_remove(pattern = "[:symbol:]") %>% as.numeric() 

dat$fib_4 <- dat$age * dat$ast_f * (dat$alt_f)^-0.5 / dat$plt_f

summary(dat$fib_4)
```

For HAV we want to look at the case criteria for HAV; including total bilitubin > 3.0 mg/dL or ALT > 200 IU/L (https://www.cdc.gov/hepatitis/hav/havfaq.htm).

```{r}
dat$bili_3 <- ifelse(lapply(dat$`BILIRUBIN TOTAL`, first) >= 3.0, 1,0)
dat$ALT_200 <- ifelse(lapply(dat$`ALT (SGPT)`, first) > 200, 1,0)
```



# Outcome Variables

## HAV Screening

IG-M is not for screening; NB HEPATITIS A AB TOTAL is an old name for HEPATITIS A IGG

```{r}
#cross-tabulating number of NA-values in both HAV IgG variables:
table(is.na(dat$`HEPATITIS A AB TOTAL`), is.na(dat$`HEPATITIS A IGG`))

dat$hav_screen <- ifelse(!(is.na(dat$`HEPATITIS A AB TOTAL`) & is.na(dat$`HEPATITIS A IGG`)),1,0)
table(dat$hav_screen, is.na(dat$`HEPATITIS A AB TOTAL`))
table(dat$hav_screen, is.na(dat$`HEPATITIS A IGG`))
table(dat$hav_screen)
```

## HAV Vaccination


```{r}
dat$hav_vaccine <- ifelse(is.na(dat$Havrix), 0, 1)
table(dat$hav_vaccine)
```

## HBV Screening

We are here using the definition of screening provided by the guidelines reviewed in Hepatitis B Online (https://www.hepatitisb.uw.edu/go/screening-diagnosis/diagnosis-hbv/core-concept/all#recommended-screening-tests); I.e. the patient needs HbSAg AND anti-HBs OR anti-HbC to be considered screened:

```{r}
# a variable for any screening test
dat$hbv_screen_any <- ifelse(
  is.na(dat$`HEP B CORE AB`) & is.na(dat$`HEP B SURFACE AB QUAL`) & is.na(dat$`HEP B SURFACE AB QUAL, SERUM`) & is.na(dat$`HEP B SURFACE AG`) & is.na(dat$`HEPATITIS B CORE AB, SERUM`) & is.na(dat$`HEPATITIS B SURFACE AG, SERUM`),
  0, 1)
# a variable for surface antigen screening
dat$hbv_sag <- ifelse(
  !is.na(dat$`HEP B SURFACE AG`) | !is.na(dat$`HEPATITIS B SURFACE AG, SERUM`),
  1, 0)
# a variable for CDC-recommended screening in PWID
dat$hbv_screen_CDC <- ifelse(
  (!is.na(dat$`HEP B SURFACE AG`) | !is.na(dat$`HEPATITIS B SURFACE AG, SERUM`)) & (!is.na(dat$`HEP B CORE AB`) | !is.na(dat$`HEP B SURFACE AB QUAL`) | !is.na(dat$`HEP B SURFACE AB QUAL, SERUM`) | !is.na(dat$`HEPATITIS B CORE AB, SERUM`)),
  1, 0)

table(dat$hbv_screen_any)
table(dat$hbv_screen_any, dat$hbv_sag)
table(dat$hbv_screen_any, dat$hbv_screen_CDC)
```

## HBV Vaccination

```{r}
dat$hbv_vaccine <- ifelse(is.na(dat$EngerixB) & is.na(dat$HeplisavB), 0, 1)
table(dat$hbv_vaccine)
```

## HDV Screening

```{r}
dat$hdv_screen <- ifelse(is.na(dat$`HEP DELTA AB`),0,1)
table(dat$hdv_screen)
```

# Misc Variables

## HBV Status

defined as positive if HBV DNA or HBSag is detected

```{r}
dat$hbv_status <- ifelse(dat$`HEP B DNA QUANT IU, VALUE` > 10 | dat$`HEP B DNA QUAL` == "Detected" | dat$`HEP B DNA QUANT IU, VALUE` > 10 | lapply(dat$`HEP B SURFACE AG`, first) == "Detected" | lapply(dat$`HEPATITIS B SURFACE AG, SERUM`, first) == "Positive" ,1,0)
dat$hbv_status[is.na(dat$hbv_status)] <- 0
table(dat$hbv_status)
```

```{r}
d2 <- dat %>%
  filter(hbv_status == 1)
head(d2)
```



# Table 1

Preliminary definitions:

```{r}
# Variables to tabulate
variables <- c("age", "genderM", "medicaid_t1", "race_eth_t1", "hiv_t1", "fib_4")

# Describes formatting of significant figures
digits_list <- list(
      age ~ 1,
      fib_4 ~ 2
    )

# Describes labels for each variable
label_list <- list(
      age ~ "Age",
      genderM ~ "Male Gender",
      race_eth_t1 ~ "Race/Ethnicity",
      medicaid_t1 ~ "Medicaid Eligible",
      hiv_t1 ~ "HIV Positive",
      fib_4 ~ "FIB-4"
    )

# Describes Statistics to compute
statistic_list <- list(
      all_continuous() ~ "{mean}",
      all_categorical() ~ "{n} ({p}%)"
    )

# Table Title
title <- "**Table 1. Patient Characteristics**"

# subset dataframe to include only data from first admissions
dat_t1 <- subset(dat, Infection_ID == first)

# Set the gtsummary theme:
theme_gtsummary_journal(journal = "jama")
```

Tabulating the sub-tables to be joined together

```{r}
t1_overall <- dat_t1 %>%
  select(variables) %>%
  tbl_summary(
    digits = digits_list,
    label = label_list,
    statistic = statistic_list,
    missing = "no"
  ) 

t1_hcv <- dat_t1 %>%
  select(hcv4, variables) %>%
  tbl_summary(
    by = hcv4,
    digits = digits_list,
    label = label_list,
    statistic = statistic_list,
    missing = "no"
  ) # %>% add_p()

t1_ivdu <- dat_t1 %>%
  select(ivdu, variables) %>%
  tbl_summary(
    by = ivdu,
    digits = digits_list,
    label = label_list,
    statistic = statistic_list,
    missing = "no"
  ) # %>% add_p() 

t1_homeless <- dat_t1 %>%
  select(homeless, variables) %>%
  tbl_summary(
    by = homeless,
    digits = digits_list,
    label = label_list,
    statistic = statistic_list,
    missing = "no"
  ) # %>% add_p()
```

Merging the tables and displaying output

```{r}
t1_merge <-
  tbl_merge(
    tbls = list(t1_overall, t1_hcv, t1_ivdu, t1_homeless),
    tab_spanner = c("**Overall**", "**HCV Status**", "**Intravenous Drug Use**", "**Housing Status**")
  ) %>%
  modify_caption(title) %>%
  bold_labels()

# cleanup
remove(digits_list, label_list, statistic_list, t1_hcv, t1_homeless, t1_ivdu, t1_overall)
```

## T1

```{r}
t1_merge
```


# Misc Calculations

## Cost of tests

### HAV IgM

HAV IGM can be considered indicated if either of the lab criteria are met (total bilitubin > 3.0 mg/dL or ALT > 200 IU/L):

```{r}
dat$hav_IgM_indicated <- ifelse(dat$bili_3 | dat$ALT_200,1,0)
table(dat$hav_IgM_indicated)
table(dat$hav_IgM_indicated, !is.na(dat$`HEPATITIS A AB, IGM`))
```

36 out of 41 HAV IgM tests ordered without lab value indication: 36*$90 = $3240.

### HBV DNA & HBV E antigen/antibody

Difficult to determine order of tests to determine if test is appropriate or not

```{r}
dat$hbv_eag <- ifelse(!is.na(dat$`HEPATITIS BE AB`) | !is.na(dat$`HEPATITIS BE AG, SERUM`), 1, 0)
table(dat$hbv_eag)
table(dat$hbv_status, dat$hbv_eag)
```

Of the 12 patients for whom the E antigen was tested, only 3 had detectable HBV DNA or HBV surface antigen; unknown cost

### HBV Core Ab

Only considered wasted if there is no surface antigen

```{r}
dat$hbv_cab <- ifelse(!is.na(dat$`HEP B COR AB IGM`) | !is.na(dat$`HEP B CORE AB`),1,0)
```


## HBV Vaccine Utilization pattern

Tabulating vaccines administered in-hospital:

```{r}
table(vaccine2$brand)
```

Results:
  * Heplisav-B (CpG 1018) was administered to 22 patients; this is notable because this vaccine was approved Nov  9, 2017, and this series requires only two shots one month apart
  * Engerix-B was administered to 14 patients
  * Havrix was administered to 17 patients (2-dose series 6-12 months apart)

## Proportion of PWHBV Screened for HCC

No data

## Proportion of PWHBV screened for HDV

```{r}
table(dat$hbv_status, dat$hdv_screen)
```

One in six were screened for HDV.

## Proportion of returning vaccinated patients receiving additiona HBV vaccine dose



## Proportion of returning unvaccinated patients starting HBV vaccine series



## Use of Accelerated Dosing

No accelerated dosing observed.

# Old Unused Code

lapply(dat$`HEP B SURFACE AB QUAL`, first) == "Detected" | dat$`HEP B SURFACE AB QUAL, SERUM` == "Reactive"

```{r}
# # Import the main dataset
# data <- read_excel(path = "data/OPTIONS DATABASE v2.0_HAV_HBV.xlsx")
# 
# # Import the extra data
# allergy <- read_excel(path = "data/HAV_HBV data_pull_de_ID.xlsx",
#                       sheet = "Allergies")
# vaccine <- read_excel(path = "data/HAV_HBV data_pull_de_ID.xlsx",
#                       sheet = "Any_vaccine")
# hep_tests <- read_excel(path = "data/HAV_HBV data_pull_de_ID.xlsx",
#                       sheet = "hepatitis_tests")
# labs <- read_excel(path = "data/HAV_HBV data_pull_de_ID.xlsx",
#                       sheet = "other_labs")
# 
# # Import chart review data
# chart <- read_excel(path = "data/STI_chart_review - deID.xlsx",
#                       sheet = "chart_review")
```